manuscript_figures

### 2024-06-20 scripts/manuscript_figures.Rmd
reran all scripts without rugosity

# Extract p-value for mean_kelp_cover
p_value_kelp_cover <- summary(lmm_model)$coefficients["mean_kelp_cover", "Pr(>|t|)"]

# Create scatter plot
figure.2 <- ggplot(aggregated_data_cover, aes(x = log(mean_kelp_cover + 1.1), y = log(mean_fish_biomass + 1.1))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_cover, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
  labs(x = "Log Kelp cover (%)", y = "Log Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Cover") +
  theme_classic();figure.2

@@

# I noticed that 06/30/23 dates were accidentily labeled 06/30/29
# Replace "29" with "23" in the DATE column
quad.filt$DATE <- sub("29$", "23", quad.filt$DATE)

@@

@@

# Extract p-value for mean_kelp_cover
p_value_kelp_biomass <- summary(lmm_model)$coefficients["mean_kelp_biomass", "Pr(>|t|)"]

# Create scatter plot
figure.3 <- ggplot(aggregated.data.morph, aes(x = log(mean_kelp_biomass + 1.1), y = log(mean_fish_biomass + 1.1))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_biomass, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
  labs(x = "Log Kelp Biomass (g/m2)", y = "Log Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Biomass") +
  theme_classic();figure.3

@@

# Fit the linear mixed effects model with log transformation
lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_rugosity + 1.1) + SITE + (1 | SITE/YEAR), data = aggregated.data.morph)

#check normality
qqPlot(residuals(lmm_model)) #improved to normal

@@

# Extract p-value for mean_kelp_cover
p_value_rugosity <- summary(lmm_model)$coefficients["mean_rugosity", "Pr(>|t|)"]

# Create scatter plot
figure.4 <- ggplot(aggregated.data.rug, aes(x = log(mean_rugosity + 1.1), y = log(mean_fish_biomass + 1.1))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
  annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_rugosity, 2)), 
           hjust = 2.1, vjust = 1.1, size = 4, parse = FALSE) +
  labs(x = "Rugosity Complexity", y = "Fish biomass (g/m2)",
       title = "Fish Biomass vs Rugosity Complexity") +
  theme_classic();figure.4

@@

# Filter quad data for all inverts and use to correlate to fish biomass
```{r}
invert.filt <-

@@

#select necessary columns
all.params.data <- master.dat %>% select("YEAR", "SITE", "COVER", "Fish.Biomass..g.m2.", "mean_rugosity", "average_kelp_biomass")

@@

# Fit a multiple regression model
fit1 <- lm(log(Fish_Biomass + 1.1) ~ log(Rugosity + 1.1) * log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1), data = all.params.data)

@@

#view results of model
summary_text1 <- capture.output(summary(fit1))
writeLines(summary_text1, "output/figures/summary_original.model.txt")

@@

# Try another type of multiple regession plot
```{r}
#fit multiple linear regression model WITH SITE AND YEAR
fit2 <- lm(Fish_Biomass ~ SITE*YEAR, data = all.params.data)

#view results of model
summary_text2 <- capture.output(summary(fit2))
writeLines(summary_text2, "output/figures/summary_model_site_year_fixed_fishbiomass.txt")

# Fit the mixed-effects model
fit3 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
              (1 | SITE) + (1 | YEAR), data = all.params.data)

#view results of model
summary_text3 <- capture.output(summary(fit3))
writeLines(summary_text3, "output/figures/summary_model_site_year_random.txt")

@@

# run with just site as random effect
# Fit the mixed-effects model
fit4 <- lmer(log(Fish_Biomass + 1.1) ~ log(Rugosity + 1.1) * log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1) + 
              (1 | SITE), data = all.params.data)

#view results of model
summary_text4 <- capture.output(summary(fit4))
writeLines(summary_text4, "output/figures/summary_model_site_random.txt")

@@

# run with just year as random effect
# Fit the mixed-effects model
fit5 <- lmer(log(Fish_Biomass + 1.1) ~ log(Rugosity + 1.1) * log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1) + 
              (1 | YEAR), data = all.params.data)

#view results of model
summary_text5 <- capture.output(summary(fit5))
writeLines(summary_text5, "output/figures/summary_model_year_random.txt")

@@

@@

# make a table for temperature values (monthly mean plus/minus one standard error) for each site (for supplement)

# Calculate monthly mean and standard error
summary_table <- temperature_data %>%
  group_by(Site, Month) %>%
  summarize(
    MeanTemp = mean(Temp),
    StdDev = sd(Temp),
    N = n(),
    StdError = StdDev / sqrt(N)
  ) %>%
  mutate(
    MeanPlusSE = MeanTemp + StdError,
    MeanMinusSE = MeanTemp - StdError
  ) %>%
  ungroup() %>%
  dplyr::select(Site, Month, MeanTemp, MeanPlusSE, MeanMinusSE)

>>>>

# Extract the p-value for the kelp biomass coefficient
p_value_kelp_cover <- lmm_summary$coefficients["log(mean_kelp_cover + 1.1)", "Pr(>|t|)"]

# Create scatter plot
figure.2 <- ggplot(aggregated_data_cover, aes(x = log(mean_kelp_cover + 1.1), y = log(mean_fish_biomass + 1.1))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +

  # R-squared values
  annotate("text", x = Inf, y = Inf, 
           label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2), 
           hjust = 1.1, vjust = 1.1, size = 4) +

  # p-value
  annotate("text", x = Inf, y = -Inf, 
           label = paste0("p = ", format(p_value_kelp_cover, scientific = FALSE, digits = 2)), 
           hjust = 1.3, vjust = -38.5, size = 4) +

  labs(x = "Log Kelp cover (%)", y = "Log Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Cover") +
  theme_classic() +
  theme(plot.margin = margin(t = 10, r = 30, b = 10, l = 10, unit = "pt")) +
  coord_cartesian(clip = 'off');figure.2

@@

# I noticed that 06/30/23 dates were accidentally labeled 06/30/29
# Replace "29" with "23" in the DATE column
quad.filt$DATE <- sub("29$", "23", quad.filt$DATE)

@@

# Identify the outliers
outliers <- c(45, 37, 25, 33)

# remove outliers
data_clean <- aggregated.data.morph[-outliers, ]

# rerun model with log and without outliers
lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_biomass + 1.1) + SITE + (1 | SITE/YEAR), data = data_clean)
#check normality
qqPlot(residuals(lmm_model)) #improved to normal

@@

# Extract the p-value for the kelp biomass coefficient
p_value_kelp_biomass <- lmm_summary$coefficients["log(mean_kelp_biomass + 1.1)", "Pr(>|t|)"]

# Create scatter plot
figure.3 <- ggplot(aggregated.data.morph, aes(x = log(mean_kelp_biomass + 1.1), y = log(mean_fish_biomass + 1.1))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +

  # R-squared values
  annotate("text", x = Inf, y = Inf, 
           label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2), 
           hjust = 1.1, vjust = 1.1, size = 4) +

  # p-value
  annotate("text", x = Inf, y = -Inf, 
           label = paste0("p = ", format(p_value_kelp_biomass, scientific = FALSE, digits = 2)), 
           hjust = 1.3, vjust = -38.5, size = 4) +

  labs(x = "Log Kelp Biomass (g/m2)", y = "Log Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Biomass") +
  theme_classic() +
  theme(plot.margin = margin(t = 10, r = 30, b = 10, l = 10, unit = "pt")) +
  coord_cartesian(clip = 'off');figure.3

@@

# Fit the linear mixed effects model with log transformation
lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_rugosity + 1.1) + SITE + (1 | SITE/YEAR), data = aggregated.data.rug)
#check normality
qqPlot(residuals(lmm_model)) #improved to normal

# Identify the outliers
outliers <- c(25, 37, 24, 35)

# remove outliers
data_clean.rug <- aggregated.data.rug[-outliers, ]

# rerun model with log and without outliers
lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_rugosity + 1.1) + SITE + (1 | SITE/YEAR), data = data_clean.rug)
#check normality
qqPlot(residuals(lmm_model)) #improved to normal

@@

# Extract the p-value for the rugosity
p_value_rugosity <- lmm_summary$coefficients["log(mean_rugosity + 1.1)", "Pr(>|t|)"]

#create scatter plot
figure.4 <- ggplot(aggregated.data.rug, aes(x = log(mean_rugosity + 1.1), y = log(mean_fish_biomass + 1.1))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +

  # R-squared values
  annotate("text", x = Inf, y = Inf, 
           label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2), 
           hjust = 1.1, vjust = 1.1, size = 4) +

  # p-value
  annotate("text", x = Inf, y = -Inf, 
           label = paste0("p = ", format(p_value_rugosity, scientific = FALSE, digits = 2)), 
           hjust = 1.3, vjust = -38.5, size = 4) +

  labs(x = "Rugosity Complexity", y = "Fish biomass (g/m2)",
       title = "Fish Biomass vs Rugosity Complexity") +
  theme_classic() +
  theme(plot.margin = margin(t = 10, r = 30, b = 10, l = 10, unit = "pt")) +
  coord_cartesian(clip = 'off');figure.4

@@

# Filter quad data for all inverts and use to correlate to fish biomass
```{r}
#invert.filt <-

@@

#select necessary columns
all.params.data <- master.dat %>% dplyr::select("YEAR", "SITE", "COVER", "Fish.Biomass..g.m2.", "mean_rugosity", "average_kelp_biomass")

@@

# Fit a multiple regression model
fit1 <- lm(log(Fish_Biomass + 1.1) ~ log(Rugosity + 1.1) * log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1), data = all.params.data)
summary(fit1)

@@

#view results of model
summary_text1 <- capture.output(summary(fit1))
writeLines(summary_text1, "output/model.result.tables/rugosity.included/summary_original.model.txt")

@@

# Try another type of multiple regession plot
```{r}
#fit multiple linear regression model WITH SITE AND YEAR
fit2 <- lm(Fish_Biomass ~ SITE*YEAR, data = all.params.data)
summary(fit2)

#view results of model
summary_text2 <- capture.output(summary(fit2))
writeLines(summary_text2, "output/model.result.tables/rugosity.included/summary_model_site_year_fixed_fishbiomass.txt")

# Fit the mixed-effects model
fit3 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
              (1 | SITE) + (1 | YEAR), data = all.params.data)
summary(fit3)

#view results of model
summary_text3 <- capture.output(summary(fit3))
writeLines(summary_text3, "output/model.result.tables/rugosity.included/summary_model_site_year_random.txt")

@@

# run with just site as random effect
# Fit the mixed-effects model
fit4 <- lmer(log(Fish_Biomass + 1.1) ~ log(Rugosity + 1.1) * log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1) + 
              (1 | SITE), data = all.params.data)
summary(fit4)

#view results of model
summary_text4 <- capture.output(summary(fit4))
writeLines(summary_text4, "output/model.result.tables/rugosity.included/summary_model_site_random.txt")

@@

# run with just year as random effect
# Fit the mixed-effects model
fit5 <- lmer(log(Fish_Biomass + 1.1) ~ log(Rugosity + 1.1) * log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1) + 
              (1 | YEAR), data = all.params.data)
summary(fit5)

#view results of model
summary_text5 <- capture.output(summary(fit5))
writeLines(summary_text5, "output/model.result.tables/rugosity.included/summary_model_year_random.txt")

@@

# Run all above models without rugosity

```{r}

# Fit a multiple regression model
fit6 <- lm(log(Fish_Biomass + 1.1) ~ log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1), data = all.params.data)
summary(fit6)

#view results of model
summary_text6 <- capture.output(summary(fit6))
writeLines(summary_text6, "output/model.result.tables/rugosity.not.included/summary_original.model.txt")

#no sig diff effects

# Fit the mixed-effects model
fit7 <- lmer(log(Fish_Biomass) ~ log(Kelp_Biomass) * log(Kelp_Cover) + (1 | SITE) + (1 | YEAR), data = all.params.data)
summary(fit7)

#view results of model
summary_text7 <- capture.output(summary(fit7))
writeLines(summary_text7, "output/model.result.tables/rugosity.not.included/summary_model_site_year_random.txt")

# run with just site as random effect
# Fit the mixed-effects model
fit8 <- lmer(log(Fish_Biomass + 1.1) ~ log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1) + (1 | SITE), data = all.params.data)
summary(fit8)

#view results of model
summary_text8 <- capture.output(summary(fit8))
writeLines(summary_text8, "output/model.result.tables/rugosity.not.included/summary_model_site_random.txt")

# site not sig different

# run with just year as random effect
# Fit the mixed-effects model
fit9 <- lmer(log(Fish_Biomass + 1.1) ~  log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1) + (1 | YEAR), data = all.params.data)
summary(fit9)

#view results of model
summary_text9 <- capture.output(summary(fit9))
writeLines(summary_text9, "output/model.result.tables/rugosity.not.included/summary_model_year_random.txt")

#year is not sig different

@@

# Make a table for temperature values (monthly mean plus/minus one standard error) for each site (for supplement)
```{r}
# Calculate monthly mean and standard error
summary_table <- temperature_data %>%
  group_by(Site, Month) %>%
  summarize(
    MeanTemp = mean(Temp),
    StdDev = sd(Temp),
    N = n(),
    StdError = StdDev / sqrt(N)
  ) %>%
  mutate(
    MeanPlusSE = MeanTemp + StdError,
    MeanMinusSE = MeanTemp - StdError
  ) %>%
  ungroup() %>%
  dplyr::select(Site, Month, MeanTemp, MeanPlusSE, MeanMinusSE)
